---
title: "Lck bulk RNA seq fig generation"
author: "Juraj Michalik"
date: "2023-10-12"
output: html_document
---

Finalized analysis for RNA seq of bulk RNA seq data of Lck. THIS IS THE FINAL VERSION!

```{r}
library(glue)
library(DESeq2)
library(GenomicFeatures)
library(GenomicAlignments)
library(tximport)
library(magrittr)
library(ggplot2)
library(pheatmap)
library(RColorBrewer)
library(gplots)
library(dplyr)
library(tidyr)
library(ggrepel)
library(Rsubread)

library(ComplexHeatmap)
library(RColorBrewer)
library(stringr)

# annotation libs
library(clusterProfiler)
library("org.Mm.eg.db")
library(AnnotationHub)

library(goseq)
library(msigdbr)
library(fgsea)
```

Function to transform inserted meta data table into ann_colors object for ComplexHeatmap annotation.

```{r}
create.annotation.colors <- function(meta.data) {
  ann.colors <- list()
  
  palette <- paste0('Set', rep(1:3, length.out = 5))
  
  i = 1
  for (col.name in colnames(meta.data)) {
    unique.levels <- unique(meta.data[[col.name]])
    num.levels <- length(unique.levels)
    
    col.pal <- colorRampPalette(brewer.pal(8, palette[i]))(num.levels) # Extend colors
    
    # Assign colors with names
    ann.colors[[col.name]] <- setNames(col.pal, unique.levels)
    i <- i + 1
  }
  
  return(ann.colors)
}
```

Function for reannotation and PCA.

```{r}
# Get current Mus Musculus database from AnnotationHub if not downloaded already
getEnsMm109 <- function(){
  aHu <- AnnotationHub()
  return(aHu[["AH109655"]])
}

# adds annotation to object
reannotate <- function(dds.obj, keyslist = row.names(dds.obj)){
  EnsMm109 <- getEnsMm109()
  symbol.vector <- mapIds(EnsMm109,
                          keys = row.names(dds.obj),
                          column = "SYMBOL",
                          keytype = "GENEID",
                          multiVals = "first")
  # rename only if symbol is a thing
  row.names(dds.obj) <- ifelse(is.na(symbol.vector) | symbol.vector == '', row.names(dds.obj), symbol.vector)
  row.names(dds.obj) <- make.unique(rownames(dds.obj))
  return(dds.obj)    
}

# returns PCA data with more PCs (12 I think)
get.complete.PCA.data <- function(dds.rlog.obj, ntop = 500){
	rvar <- rowVars(assay(dds.rlog.obj))
	select <- order(rvar, decreasing=TRUE)[seq_len(min(ntop, length(rvar)))]
	pca.dat <- prcomp(t(assay(dds.rlog.obj)[select,]))
	return(pca.dat)
}
```

Function to make GSEA plot. Returns ggplot object.

```{r}
# Helper function to help plot data
printGSEA_pars <- function(title, GSEA_res, row_num){
  printed.str <- paste0(title, '\n', '(pval: ', formatC(GSEA_res[row_num,2] %>% as.numeric(), format = "E", digits = 2) , 
                        ', ES: ', round(GSEA_res[row_num, 5],3), 
                        ', NES: ', round(GSEA_res[row_num, 6],3), ', size: ', GSEA_res[row_num, 7], ')')
  return(printed.str)
}

# Plots GSEA
# - GSE_data: data to plot, export from plotEnrichment function
# - title: title of plot
# - GSEA_res: results to show in title
# - row_num: row of GSEA results to show results of (in title)
# - title.size: size of letters for title
# Returns gsea object
plotGSEA <- function(GSE_data, title, GSEA_res, row_num, title.size = 14){
  # Calculate intersects - required for area computations
  GSE_data.area51 <- GSE_data %>% as.data.frame
  i <- nrow(GSE_data.area51)
  while(i > 1){
    if(GSE_data.area51$ES[i-1] * GSE_data.area51$ES[i] < 0){
      ratio <- abs(GSE_data.area51$ES[i-1])/(abs(GSE_data.area51$ES[i-1]) + abs(GSE_data.area51$ES[i]))
      rank.cross <- GSE_data.area51$rank[i-1] + (GSE_data.area51$rank[i] - GSE_data.area51$rank[i-1]) * ratio
      GSE_data.area51 <- tibble::add_row(GSE_data.area51, rank = rank.cross, ES = 0, .before = i)
    }
    i <- i - 1
  }
  
  # plot
  ggplot.obj <- ggplot(GSE_data, aes(x = rank, y = ES))+
    geom_line(color = "#2E86AB", linewidth = 1.2) +
    geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
    geom_hline(aes(yintercept = ifelse(abs(max(ES)) > abs(min(ES)), max(ES), min(ES))), color = "firebrick3", linetype = 'dashed') +
    geom_area(data = GSE_data.area51, mapping = aes(x = rank, y = ifelse(ES > 0, ES, 0)), fill = "#ACADFF", alpha = 0.3) +
    geom_area(data = GSE_data.area51, mapping = aes(x = rank, y = ifelse(ES < 0, ES, 0)), fill = "firebrick1", alpha = 0.3) +
    geom_segment(
      aes(x = rank,
          xend = rank,
          y = -max(abs(ES))*0.05,
          yend = max(abs(ES))*0.05),
      linewidth = 0.3
    ) + theme_classic() +
    ggtitle(printGSEA_pars(title, GSEA_res, row_num)) +
    ylab('Enrichment Score (ES)') +
    theme(axis.title = element_text(face = "bold"),
          plot.title = element_text(hjust = 0.5, face = 'bold', size = title.size))
  
  return(ggplot.obj)
}
```

Finally we setup paths to directories and files, plus some statistics.
Bamfiles were generated by STAR 2.7.10, reference with the same software from the gtf file below and GRCm39 assembly.

```{r}
page.title <- "Lck data analysis - combined run"
bam.path <- "BAM/"
gtf.path <- "Mus_musculus.GRCm39.109.gtf"
sample.path <- "Sample_sheet.csv"

# filtering statistic
padj.max <- 0.01
lfc.min <- 0.6
```

Create directories where figures and data will be stored.

```{r}
dir.create('Main_figures', showWarnings = FALSE)
dir.create('Supplemental_figures', showWarnings = FALSE)
dir.create('Pathways/Filtered_any', showWarnings = FALSE)
dir.create('Images', showWarnings = FALSE)
dir.create('Data', showWarnings = FALSE)
```

# Building count table

```{r}
files <- list.files(bam.path, pattern = "\\.bam$")
files <- paste0(bam.path, files) %>% gtools::mixedsort()

# get reference and set it up
(txdb <- makeTxDbFromGFF(gtf.path, format = "gtf", circ_seqs = character()))
ebg <- exonsBy(txdb, by = "gene")

# quantification step
expcounts <- summarizeOverlaps(features = ebg,
                               reads = files,
                               mode = "Union",
                               ignore.strand = FALSE,
                               preprocess.reads = invertStrand,
                               singleEnd = FALSE, 
                               fragments = TRUE)
```

```{r}
# remove bam from colnames
colnames(expcounts) <- gsub('Aligned.sortedByCoord.out.bam', '', colnames(expcounts))
```

Load sample tab. Additionally, we need to create combined parameter, as we'll need to compute contrasts for those.

```{r}
sample_tab <- read.csv(sample.path) %>% dplyr::mutate(genotype.act.inh = paste(genotype, activation, inhibition))
```

Reorder samples in expcounts to match those in sample tab.

```{r}
idx.reord <- match(sample_tab$sample.name, colnames(expcounts))

expcounts <- expcounts[,idx.reord]
```

Verify.

```{r}
# Consistency check, must be TRUE
identical(as.character(sample_tab$sample.name), colnames(expcounts))

# Should be already ordered by mixedsort
rownames(sample_tab) <- as.character(sample_tab$sample.name)
colData(expcounts) <- DataFrame(sample_tab)
```

```{r}
round(colSums(assay(expcounts)) / 1e6, 1)
```

# Preparation of DESeq2 object

Here we prepare DESeq object. We want to take into account KO factor but also activation. Finally, there may be differences between Experiments. Design must account for all of these variables. We also remove low-count, noisy genes and reannotate data.

```{r}
dds_bulk_Lck.full_data <- DESeqDataSet(expcounts, design = ~ exp.num + genotype.act.inh)
rowData(dds_bulk_Lck.full_data)$EnsemblID <- rownames(expcounts)
dds_bulk_Lck.full_data <- dds_bulk_Lck.full_data[rowSums(counts(dds_bulk_Lck.full_data)) > 10, ] # remove <= 10 counts = noise
dds_bulk_Lck.full_data <- reannotate(dds_bulk_Lck.full_data)
```

```{r}
dds_bulk_Lck.full_data <- DESeq(dds_bulk_Lck.full_data)
```

Save the object.

```{r}
saveRDS(dds_bulk_Lck.full_data, 'Data/Lck_dds_full.rds')
```

Write also full assay as CSV file.

```{r}
write.csv(assay(dds_bulk_Lck.full_data), 'Lck_all_samples_raw_counts_annotated.csv')
```

# QC analysis

We perform blinded analysis that is not dependent from design.

```{r}
rlog_bulk.blinded.full_data <- rlog(dds_bulk_Lck.full_data, blind = TRUE)
```

## PCA

Compute and prepare blinded PCA.

```{r}
blinded_pca_full_data_cmp <- get.complete.PCA.data(rlog_bulk.blinded.full_data)
blinded_pca_full_data <- blinded_pca_full_data_cmp$x %>% as.data.frame %>% cbind(., colData(dds_bulk_Lck.full_data)) %>% 
  dplyr::mutate(act.inhibitor = paste(activation, inhibition, sep = '-'),
                act.inhibitor = gsub('-None','', act.inhibitor),
                exp.desc = paste('Exp', exp.num))

                
# get fraction of variability captured by each PC
blinded_pca_full_data_Var <- blinded_pca_full_data_cmp$sdev^2/sum(blinded_pca_full_data_cmp$sdev^2)
```

```{r}
blinded.pca12.full_data.base <- ggplot(blinded_pca_full_data, aes(x = PC1, y = PC2, color = genotype, shape = act.inhibitor)) +
  geom_point(size = 5) +
  labs(color = "activator/inhibitor") +  
  scale_color_manual(values = c('firebrick1', 'forestgreen')) +
  ggtitle(paste0('LCK bulk RNAseq - Normalized PCA (design-blind, PC1 + PC2)' )) +
  xlab(paste0("PC1: ", round(blinded_pca_full_data_Var[1]*100), "% variance")) +
  ylab(paste0("PC2: ", round(blinded_pca_full_data_Var[2]*100), "% variance")) +
  theme_classic()

blinded.pca23.full_data.base <- ggplot(blinded_pca_full_data, aes(x = PC2, y = PC3, color = genotype, shape = act.inhibitor)) +
  geom_point(size = 5) +
  labs(color = "activator/inhibitor") +
  scale_color_manual(values = c('firebrick1', 'forestgreen')) +
  ggtitle(paste0('LCK bulk RNAseq - Normalized PCA (design-blind, PC2 + PC3)' )) +
  xlab(paste0("PC2: ", round(blinded_pca_full_data_Var[2]*100), "% variance")) +
  ylab(paste0("PC3: ", round(blinded_pca_full_data_Var[3]*100), "% variance")) +
  theme_classic()

blinded.pca12.full_data <- blinded.pca12.full_data.base +
  geom_text_repel(aes(label = paste(exp.desc)), seed = 120)

#pca12.blank <- pca12.base +
#  geom_text_repel(aes(label = Sample), seed = 120, color = NA, segment.color = 'white') +
#  theme(
#    text = element_text(color = "white"),
#    legend.text = element_text(color = "white"),
#    legend.title = element_text(color = "white"),
#    strip.text = element_text(color = "white"),
#    axis.text.x = element_text(color = "white"),
#    axis.text.y = element_text(color = "white"))

blinded.pca23.full_data <- blinded.pca23.full_data.base +
  geom_text_repel(aes(label = paste(exp.desc)), seed = 120)

#pca23.blank <- pca23.base +
#  geom_text_repel(aes(label = Sample), seed = 120, color = NA, segment.color = 'white') +
#  theme(
#    text = element_text(color = "white"),
#    legend.text = element_text(color = "white"),
#    legend.title = element_text(color = "white"),
#    strip.text = element_text(color = "white"),
#    axis.text.x = element_text(color = "white"),
#    axis.text.y = element_text(color = "white"))

print(blinded.pca12.full_data)
print(blinded.pca23.full_data)
```

```{r}
tiff('Images/Lck_bulk_all_full_PCA_blinded_PC1_PC2.tiff', width = 700, height = 500, res = 100)
print(blinded.pca12.full_data)
dev.off()

tiff('Images/Lck_bulk_PCA_all_full_blinded_PC2_PC3.tiff', width = 700, height = 500, res = 100)
print(blinded.pca23.full_data)
dev.off()

```

# Save data

We calculate normalized count table.

```{r}
Lck_normalized_full <- counts(dds_bulk_Lck.full_data, normalized = TRUE)
```

Write it also to file, along with raw data.

```{r}
write.csv(counts(dds_bulk_Lck.full_data), 'Data/Lck_all_runs_table_full.csv')
write.csv(Lck_normalized_full, 'Data/Lck_all_runs_bulk_normalized_table_full.csv')
```

# Preparation of DESeq object from exp2 and exp3 and gene filtering

Now we make DESeq object from exp2 and exp3 only. The data will be made from filtered count matrix, where genes will be kept according to the filtering, and exp1 will be thrown away. We then filter sequences out.

```{r}
expcounts.subs <- expcounts[,6:25]
```

Make DDS object.

```{r}
dds_bulk_Lck.data <- DESeqDataSet(expcounts.subs, design = ~ exp.num + genotype.act.inh)
rowData(dds_bulk_Lck.data)$EnsemblID <- rownames(expcounts.subs)
dds_bulk_Lck.data <- dds_bulk_Lck.data[rowSums(counts(dds_bulk_Lck.data)) > 10, ] # remove <= 10 counts = noise
dds_bulk_Lck.data <- reannotate(dds_bulk_Lck.data)
```

Run DESeq, to re-normalize.

```{r}
dds_bulk_Lck.data <- DESeq(dds_bulk_Lck.data)
```

Next we will compute the list of pathways that we'll test enrichment for. First, we'll pre-filter normalized count table so at least two samples from exp2 and at least 2 samples from exp3 have higher normalized count than 16.

```{r}
exp.num <- colData(dds_bulk_Lck.data)$exp.num
norm.counts <- counts(dds_bulk_Lck.data, normalized = TRUE)

keep <- rowSums(norm.counts[, exp.num == 2, drop = FALSE] >= 16) >= 2 &
        rowSums(norm.counts[, exp.num == 3, drop = FALSE] >= 16) >= 2

dds_bulk_Lck.data <- dds_bulk_Lck.data[keep, ]
```

Rerun DESeq on filtered data.

```{r}
dds_bulk_Lck.data <- DESeq(dds_bulk_Lck.data)
```

Save the object.

```{r}
saveRDS(dds_bulk_Lck.data, 'Data/Lck_dds_exp2_exp3.rds')
```

Get rlog values.

```{r}
rlog_bulk.blinded.data <- rlog(dds_bulk_Lck.data, blind = TRUE)
```

## PCA

We show PCA again, this time for exp2 and3 only.

```{r}
blinded_pca_data_cmp <- get.complete.PCA.data(rlog_bulk.blinded.data)
blinded_pca_data <- blinded_pca_data_cmp$x %>% as.data.frame %>% cbind(., colData(dds_bulk_Lck.data)) %>% 
  dplyr::mutate(act.inhibitor = paste(activation, inhibition, sep = '-'),
                act.inhibitor = gsub('-None','', act.inhibitor),
                exp.desc = paste('Exp', exp.num))

# get fraction of variability captured by each PC
blinded_pca_Var <- blinded_pca_data_cmp$sdev^2/sum(blinded_pca_data_cmp$sdev^2)
```

```{r}
blinded.pca12.base <- ggplot(blinded_pca_data, aes(x = PC1, y = PC2, color = genotype, shape = act.inhibitor)) +
  geom_point(size = 5) +
  labs(color = "activator/inhibitor") +  
  scale_color_manual(values = c('firebrick1', 'forestgreen')) +
  ggtitle(paste0('LCK bulk RNAseq - Normalized PCA (design-blind, PC1 + PC2)' )) +
  xlab(paste0("PC1: ", round(blinded_pca_Var[1]*100), "% variance")) +
  ylab(paste0("PC2: ", round(blinded_pca_Var[2]*100), "% variance")) +
  theme_classic()

blinded.pca23.base <- ggplot(blinded_pca_data, aes(x = PC2, y = PC3, color = genotype, shape = act.inhibitor)) +
  geom_point(size = 5) +
  labs(color = "activator/inhibitor") +
  scale_color_manual(values = c('firebrick1', 'forestgreen')) +
  ggtitle(paste0('LCK bulk RNAseq - Normalized PCA (design-blind, PC2 + PC3)' )) +
  xlab(paste0("PC2: ", round(blinded_pca_Var[2]*100), "% variance")) +
  ylab(paste0("PC3: ", round(blinded_pca_Var[3]*100), "% variance")) +
  theme_classic()

blinded.pca12 <- blinded.pca12.base +
  geom_text_repel(aes(label = paste(exp.desc)), seed = 120)

#pca12.blank <- pca12.base +
#  geom_text_repel(aes(label = Sample), seed = 120, color = NA, segment.color = 'white') +
#  theme(
#    text = element_text(color = "white"),
#    legend.text = element_text(color = "white"),
#    legend.title = element_text(color = "white"),
#    strip.text = element_text(color = "white"),
#    axis.text.x = element_text(color = "white"),
#    axis.text.y = element_text(color = "white"))

blinded.pca23 <- blinded.pca23.base +
  geom_text_repel(aes(label = paste(exp.desc)), seed = 120)

#pca23.blank <- pca23.base +
#  geom_text_repel(aes(label = Sample), seed = 120, color = NA, segment.color = 'white') +
#  theme(
#    text = element_text(color = "white"),
#    legend.text = element_text(color = "white"),
#    legend.title = element_text(color = "white"),
#    strip.text = element_text(color = "white"),
#    axis.text.x = element_text(color = "white"),
#    axis.text.y = element_text(color = "white"))

print(blinded.pca12)
print(blinded.pca23)
```

# Save data

We calculate normalized count table.

```{r}
Lck_normalized <- counts(dds_bulk_Lck.data, normalized = TRUE)
```

Write it also to file, along with raw data.

```{r}
write.csv(counts(dds_bulk_Lck.data), 'Data/Lck_all_runs_table_exp2_exp3.csv')
write.csv(Lck_normalized, 'Data/Lck_all_runs_bulk_normalized_table_exp2_exp3.csv')
```

# Preparation of pathways

Next we will compute the list of pathways that we'll test enrichment for. First, we pre-filter data to values based on genes filtered from exp2 and exp3-containing data.

```{r}
Lck_normalized_full_filt <- Lck_normalized_full[rownames(Lck_normalized_full) %in% rownames(dds_bulk_Lck.data),]
```

Now we keep only samples from first experiment, that will be used for creation of pathway lists. We also calculate log2(x + 1) (pseudocount) values.

```{r}
CountTable_filt <- Lck_normalized_full_filt[,c(1:5)]
CountTable_log2 <- log2(CountTable_filt + 1)
rownames(CountTable_log2) <- rownames(CountTable_log2)

colnames(CountTable_log2) <- c("NA", "OVA", "OVA+p38i", "OVA+Erki", "OVA+NFATi")
```

Prepare function to filter genes.

```{r}
filter_gene_list <- function(list_name, contrast_col, direction, suffix_label, fc1.2 = log2(1.2)) {
  genes <- read.csv(
    paste0("Pathways/Unfiltered/", list_name, ".csv"),
    header = FALSE,
    stringsAsFactors = FALSE
  )[[1]]
  
  genes_present <- unique(genes[genes %in% rownames(CountTable_log2)])
  
  tab <- data.frame(
    Gene = genes_present,
    OVA = CountTable_log2[genes_present, "OVA"],
    Inhibitor = CountTable_log2[genes_present, contrast_col],
    Delta = CountTable_log2[genes_present, contrast_col] - CountTable_log2[genes_present, "OVA"],
    row.names = NULL
  )
  
  if (direction == "down") {
    filtered_any <- tab$Gene[tab$Delta < 0]
    filtered_1.2x <- tab$Gene[tab$Delta <= -fc1.2]
    cat("Genes down in", contrast_col, "vs OVA:", length(filtered_any), "\n")
    cat("Genes down at least 1.2-fold:", length(filtered_1.2x), "\n")
  } else if (direction == "up") {
    filtered_any <- tab$Gene[tab$Delta > 0]
    filtered_1.2x <- tab$Gene[tab$Delta >= fc1.2]
    cat("Genes up in", contrast_col, "vs OVA:", length(filtered_any), "\n")
    cat("Genes up at least 1.2-fold:", length(filtered_1.2x), "\n")
  } else {
    stop("direction must be 'down' or 'up'")
  }
  
  cat("\n---", list_name, "---\n")
  cat("Genes in input list:", length(genes), "\n")
  cat("Genes found in table:", length(genes_present), "\n")
  
  write.table(
    filtered_any,
    paste0("Pathways/Filtered_any/", list_name, "_filtered_any", suffix_label, ".csv"),
    sep = ",",
    row.names = FALSE,
    col.names = FALSE,
    quote = FALSE
  )
  
  write.table(
    filtered_1.2x,
    paste0("Pathways/Filtered/", list_name, "_filtered_1.2x.csv"),
    sep = ",",
    row.names = FALSE,
    col.names = FALSE,
    quote = FALSE
  )
  
  return(tab)
}
```

Use the above and table of genes to generate gene lists.

```{r}
NFATUP_table <- filter_gene_list("NFATUP", "OVA+NFATi", "down", "UP")
NFATDOWN_table <- filter_gene_list("NFATDOWN", "OVA+NFATi", "up", "DOWN")

ERKUP_table <- filter_gene_list("ERKUP", "OVA+Erki", "down", "UP")
ERKDOWN_table <- filter_gene_list("ERKDOWN", "OVA+Erki", "up", "DOWN")

P38UP_table <- filter_gene_list("P38UP", "OVA+p38i", "down", "UP")
P38DOWN_table <- filter_gene_list("P38DOWN", "OVA+p38i", "up", "DOWN")

# Validation
summary(ERKUP_table$Delta)
head(ERKUP_table[order(ERKUP_table$Delta),])
tail(ERKUP_table[order(ERKUP_table$Delta),])
```

Generate top 250 Activated/Inhibited gene pathway.

```{r}
delta_OVA <- CountTable_log2[, "OVA"] - CountTable_log2[, "NA"]
delta_OVA <- sort(delta_OVA[!is.na(delta_OVA)], decreasing = TRUE)

TOP250activatedOVA <- names(delta_OVA)[1:250]
TOP250inhibitedOVA <- names(delta_OVA)[(length(delta_OVA)-249):length(delta_OVA)]

#Checks
length(TOP250activatedOVA); length(TOP250inhibitedOVA)

write.table(TOP250activatedOVA,  "Pathways/Filtered/TOP250activatedOVA.csv",  sep = ",", row.names = FALSE, col.names = FALSE, quote = FALSE)
write.table(TOP250inhibitedOVA,  "Pathways/Filtered/TOP250inhibitedOVA.csv",  sep = ",", row.names = FALSE, col.names = FALSE, quote = FALSE)
```

# Heatmap scores generation

We now plot heatmaps of combined samples from data. For this purpose, we take count data from matrix that only includes exp2 and exp3. We then apply log2(x + 1) transform on it.

```{r}
CountTable_log2 <- log2(Lck_normalized + 1)
rownames(CountTable_log2) <- rownames(Lck_normalized)
```

Reorder data to have same samples from both experiments together

```{r}
CountTable_log2EXP <- CountTable_log2[, c(
  "Sample6Bulk1",  "Sample1Bulk2", "Sample11Bulk1", "Sample6Bulk2",
  "Sample7Bulk1",  "Sample2Bulk2", "Sample12Bulk1", "Sample7Bulk2",
  "Sample8Bulk1",  "Sample3Bulk2", "Sample13Bulk1", "Sample8Bulk2",
  "Sample9Bulk1",  "Sample4Bulk2", "Sample14Bulk1", "Sample9Bulk2",
  "Sample10Bulk1", "Sample5Bulk2", "Sample15Bulk1", "Sample10Bulk2"
)]

colnames(CountTable_log2EXP) <- c(
  "WT-NA-2",
  "WT-NA-3",
  "KO-NA-2",
  "KO-NA-3", 
  "WT-OVA-2",
  "WT-OVA-3",  
  "KO-OVA-2",
  "KO-OVA-3",  
  "WT-p38i-2",
  "WT-p38i-3",  
  "KO-p38i-2",
  "KO-p38i-3",  
  "WT-ERKi-2",
  "WT-ERKi-3",  
  "KO-ERKi-2",
  "KO-ERKi-3",  
  "WT-NFATi-2",
  "WT-NFATi-3",  
  "KO-NFATi-2",
  "KO-NFATi-3"  
)
```

Load gene list files we previously created.

```{r}
folder <- "Pathways/Filtered/"

# gene list files - edit this vector to your actual filenames
gene_list_files <- c(
  "TOP250activatedOVA.csv",
  "P38UP_filtered_1.2x.csv",
  "ERKUP_filtered_1.2x.csv",
  "NFATUP_filtered_1.2x.csv",
  "NFKBall.csv",
  "TOP250inhibitedOVA.csv",
  "P38DOWN_filtered_1.2x.csv",
  "ERKDOWN_filtered_1.2x.csv",
  "NFATDOWN_filtered_1.2x.csv"
)
```

We now load the function that will be used to calculate values, that will be then plotted to heatmap.

```{r}
# function to calculate one gene-list score across samples
calc_gene_score <- function(file, expr_mat, do.scaling = TRUE) {
  gene_list <- read.csv(
    file.path(folder, file),
    header = FALSE,
    stringsAsFactors = FALSE
  )[[1]]
  
  gene_list <- unique(as.character(gene_list))
  genes_present <- gene_list[gene_list %in% rownames(expr_mat)]
  
  if (length(genes_present) == 0) {
    score <- rep(NA, ncol(expr_mat))
    names(score) <- colnames(expr_mat)
    return(score)
  }
  
  mat <- expr_mat[genes_present, , drop = FALSE]
  
  # remove genes with zero variance
  mat <- mat[apply(mat, 1, sd, na.rm = TRUE) > 0, , drop = FALSE]
  
  if (nrow(mat) == 0) {
    score <- rep(NA, ncol(expr_mat))
    names(score) <- colnames(expr_mat)
    return(score)
  }
  
  mat_z <- mat
  if(do.scaling == TRUE){ # row-wise z-score if scaling to apply
    print('Scaling enabled: scaling now.')
    mat_z <- t(scale(t(mat)))
  }
  
  # average z-score per sample
  score <- colMeans(mat_z, na.rm = TRUE)
  return(score)
}
```

Calculate scores.

```{r}
# calculate scores for all gene lists
ScoreTable <- do.call(
  rbind,
  lapply(gene_list_files, function(f) calc_gene_score(f, CountTable_log2EXP))
)

rownames(ScoreTable) <- tools::file_path_sans_ext(basename(gene_list_files))
```

Save the score table.

```{r}
write.csv(ScoreTable, file.path(folder, "GeneList_ScoreTable.csv"), quote = FALSE)
```

Plot heatmaps.

```{r}
#HEAT MAP WITH ALL 20 SAMPLES
pheatmap(
  as.matrix(ScoreTable),
  scale = "row",
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  color = colorRampPalette(c("blue", "white", "red"))(100)
)

#HEAT MAP WITH 16 SAMPLES WHICH WERE OVA-ACTIVATED
pheatmap(
  as.matrix(ScoreTable[, -c(1:4)]),
  scale = "row",
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  color = colorRampPalette(c("blue", "white", "red"))(100)
)

#WITHOUT inhibitors
pheatmap(
  as.matrix(ScoreTable[, c(1:8)]),
  scale = "row",
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  color = colorRampPalette(c("blue", "white", "red"))(100)
)

#ONLY SAMPLES WITHOUT Inhibitors - COLOR SCALING
mat <- as.matrix(ScoreTable[, c(1:8)])

my_breaks <- c(
  seq(-2, 0, length.out = 20),
  seq(0.001, 1.5, length.out = 80)
)

#HEAT MAP WITH 16 SAMPLES WHICH WERE OVA-ACTIVATED
pheatmap(
  as.matrix(ScoreTable[, -c(1:4)]),
  scale = "row",
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  color = colorRampPalette(c("blue", "white", "red"))(100),
  breaks = my_breaks
)

pheatmap(
  mat,
  scale = "row",
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  color = colorRampPalette(c("blue", "white", "red"))(100),
  breaks = my_breaks
)
```

# Differential expression analysis and GSEA

We now perform differential expression analyses and generate results.

```{r}
print(dds_bulk_Lck.data)
print(dds_bulk_Lck.data@design)
print(dds_bulk_Lck.data@colData)
```

Now get results. No LFC shrinkage will be applied.

```{r}
res.WT.vs.KO.OVA <- results(
  dds_bulk_Lck.data,
  contrast = c("genotype.act.inh", "WT OVA 1uM None", "KO OVA 1uM None"),
)

##

res.WT.vs.WT.IP38 <- results(
  dds_bulk_Lck.data,
  contrast = c("genotype.act.inh", "WT OVA 1uM None", "WT OVA 1uM Ip38 12.5 uM"),
)

res.WT.vs.WT.ERK <- results(
  dds_bulk_Lck.data,
  contrast = c("genotype.act.inh", "WT OVA 1uM None", "WT OVA 1uM inhERK10uM"),
)

res.WT.vs.WT.NFAT <- results(
  dds_bulk_Lck.data,
  contrast = c("genotype.act.inh", "WT OVA 1uM None", "WT OVA 1uM Cyclosporin A 0.1uM"),
)

##

res.KO.vs.KO.IP38 <- results(
  dds_bulk_Lck.data,
  contrast = c("genotype.act.inh", "KO OVA 1uM None", "KO OVA 1uM Ip38 12.5 uM"),
)

res.KO.vs.KO.ERK <- results(
  dds_bulk_Lck.data,
  contrast = c("genotype.act.inh", "KO OVA 1uM None", "KO OVA 1uM inhERK10uM"),
)

res.KO.vs.KO.NFAT <- results(
  dds_bulk_Lck.data,
  contrast = c("genotype.act.inh", "KO OVA 1uM None", "KO OVA 1uM Cyclosporin A 0.1uM"),
)

##

res.WT.IP38.vs.KO.IP38 <- results(
  dds_bulk_Lck.data,
  contrast = c("genotype.act.inh", "WT OVA 1uM Ip38 12.5 uM", "KO OVA 1uM Ip38 12.5 uM"),
)

res.WT.ERK.vs.KO.ERK <- results(
  dds_bulk_Lck.data,
  contrast = c("genotype.act.inh", "WT OVA 1uM inhERK10uM", "KO OVA 1uM inhERK10uM"),
)

res.WT.NFAT.vs.KO.NFAT <- results(
  dds_bulk_Lck.data,
  contrast = c("genotype.act.inh", "WT OVA 1uM Cyclosporin A 0.1uM", "KO OVA 1uM Cyclosporin A 0.1uM"),
)

```

Create fold-change lists (not shrunk) and sort them.

```{r}
lfc.res.WT.vs.KO.OVA <- res.WT.vs.KO.OVA$log2FoldChange
names(lfc.res.WT.vs.KO.OVA) <- rownames(res.WT.vs.KO.OVA)
lfc.res.WT.vs.KO.OVA <- sort(lfc.res.WT.vs.KO.OVA, decreasing = T)

##

lfc.res.WT.vs.WT.IP38 <- res.WT.vs.WT.IP38$log2FoldChange
names(lfc.res.WT.vs.WT.IP38) <- rownames(res.WT.vs.WT.IP38)
lfc.res.WT.vs.WT.IP38 <- sort(lfc.res.WT.vs.WT.IP38, decreasing = T)

lfc.res.WT.vs.WT.ERK <- res.WT.vs.WT.ERK$log2FoldChange
names(lfc.res.WT.vs.WT.ERK) <- rownames(res.WT.vs.WT.ERK)
lfc.res.WT.vs.WT.ERK <- sort(lfc.res.WT.vs.WT.ERK, decreasing = T)

lfc.res.WT.vs.WT.NFAT <- res.WT.vs.WT.NFAT$log2FoldChange
names(lfc.res.WT.vs.WT.NFAT) <- rownames(res.WT.vs.WT.NFAT)
lfc.res.WT.vs.WT.NFAT <- sort(lfc.res.WT.vs.WT.NFAT, decreasing = T)

##

lfc.res.KO.vs.KO.IP38 <- res.KO.vs.KO.IP38$log2FoldChange
names(lfc.res.KO.vs.KO.IP38) <- rownames(res.KO.vs.KO.IP38)
lfc.res.KO.vs.KO.IP38 <- sort(lfc.res.KO.vs.KO.IP38, decreasing = T)

lfc.res.KO.vs.KO.ERK <- res.KO.vs.KO.ERK$log2FoldChange
names(lfc.res.KO.vs.KO.ERK) <- rownames(res.KO.vs.KO.ERK)
lfc.res.KO.vs.KO.ERK <- sort(lfc.res.KO.vs.KO.ERK, decreasing = T)

lfc.res.KO.vs.KO.NFAT <- res.KO.vs.KO.NFAT$log2FoldChange
names(lfc.res.KO.vs.KO.NFAT) <- rownames(res.KO.vs.KO.NFAT)
lfc.res.KO.vs.KO.NFAT <- sort(lfc.res.KO.vs.KO.NFAT, decreasing = T)

##

lfc.res.WT.IP38.vs.KO.IP38 <- res.WT.IP38.vs.KO.IP38$log2FoldChange
names(lfc.res.WT.IP38.vs.KO.IP38) <- rownames(res.WT.IP38.vs.KO.IP38)
lfc.res.WT.IP38.vs.KO.IP38 <- sort(lfc.res.WT.IP38.vs.KO.IP38, decreasing = T)

lfc.res.WT.ERK.vs.KO.ERK <- res.WT.ERK.vs.KO.ERK$log2FoldChange
names(lfc.res.WT.ERK.vs.KO.ERK) <- rownames(res.WT.ERK.vs.KO.ERK)
lfc.res.WT.ERK.vs.KO.ERK <- sort(lfc.res.WT.ERK.vs.KO.ERK, decreasing = T)

lfc.res.WT.NFAT.vs.KO.NFAT <- res.WT.NFAT.vs.KO.NFAT$log2FoldChange
names(lfc.res.WT.NFAT.vs.KO.NFAT) <- rownames(res.WT.NFAT.vs.KO.NFAT)
lfc.res.WT.NFAT.vs.KO.NFAT <- sort(lfc.res.WT.NFAT.vs.KO.NFAT, decreasing = T)
```

Set up pathway list as a single object.

```{r}
pathways_list_full <- list()
top250act.up <- read.csv('Pathways/Filtered/TOP250activatedOVA.csv', header = F)[,1]
top250act.down <- read.csv('Pathways/Filtered/TOP250inhibitedOVA.csv', header = F)[,1]
ERK_down <- read.csv('Pathways/Filtered/ERKDOWN_filtered_1.2x.csv', header = F)[,1]
ERK_up <- read.csv('Pathways/Filtered/ERKUP_filtered_1.2x.csv', header = F)[,1]
NFAT_down <- read.csv('Pathways/Filtered/NFATDOWN_filtered_1.2x.csv', header = F)[,1]
NFAT_up <- read.csv('Pathways/Filtered/NFATUP_filtered_1.2x.csv', header = F)[,1]
P38_down <- read.csv('Pathways/Filtered/P38DOWN_filtered_1.2x.csv', header = F)[,1]
P38_up <- read.csv('Pathways/Filtered/P38UP_filtered_1.2x.csv', header = F)[,1]
NFKBall <- read.csv('Pathways/Filtered/NFKBall.csv', header = F)[,1]

pathways_list_full[['top250act.up']] <- top250act.up
pathways_list_full[['top250act.down']] <- top250act.down
pathways_list_full[['ERK_down']] <- ERK_down
pathways_list_full[['ERK_up']] <- ERK_up
pathways_list_full[['NFAT_down']] <- NFAT_down
pathways_list_full[['NFAT_up']] <- NFAT_up
pathways_list_full[['P38_down']] <- P38_down
pathways_list_full[['P38_up']] <- P38_up
pathways_list_full[['NFKB_all']] <- NFKBall

pathways_list <- pathways_list_full
```

Calculate GSEA for each pathway and contrast.

```{r}
set.seed(42)
fgsea_Lck_lfc.res.WT.vs.KO.OVA <- fgsea(pathways = pathways_list,
                                       stats = lfc.res.WT.vs.KO.OVA)

##

fgsea_Lck_lfc.res.WT.vs.WT.IP38 <- fgsea(pathways = pathways_list,
                                       stats = lfc.res.WT.vs.WT.IP38)

fgsea_Lck_lfc.res.WT.vs.WT.ERK <- fgsea(pathways = pathways_list,
                                       stats = lfc.res.WT.vs.WT.ERK)

fgsea_Lck_lfc.res.WT.vs.WT.NFAT <- fgsea(pathways = pathways_list,
                                         stats = lfc.res.WT.vs.WT.NFAT)

##

fgsea_Lck_lfc.res.KO.vs.KO.IP38 <- fgsea(pathways = pathways_list,
                                       stats = lfc.res.KO.vs.KO.IP38)

fgsea_Lck_lfc.res.KO.vs.KO.ERK <- fgsea(pathways = pathways_list,
                                        stats = lfc.res.KO.vs.KO.ERK)

fgsea_Lck_lfc.res.KO.vs.KO.NFAT <- fgsea(pathways = pathways_list,
                                         stats = lfc.res.KO.vs.KO.NFAT)

##

fgsea_Lck_lfc.res.WT.IP38.vs.KO.IP38 <- fgsea(pathways = pathways_list,
                                              stats = lfc.res.WT.IP38.vs.KO.IP38)

fgsea_Lck_lfc.res.WT.ERK.vs.KO.ERK <- fgsea(pathways = pathways_list,
                                              stats = lfc.res.WT.ERK.vs.KO.ERK)

fgsea_Lck_lfc.res.WT.NFAT.vs.KO.NFAT <- fgsea(pathways = pathways_list,
                                              stats = lfc.res.WT.NFAT.vs.KO.NFAT)

```

## GSEA plots

Plot all GSEAs.

## WT vs. KO - act - up

GSEA WT/KO Act. contrast (Activation-induced genes)

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list_full[['top250act.up']],  lfc.res.WT.vs.KO.OVA)$data
GSEA.WTKO_act_250_top <- plotGSEA(GSE_data, 'Gene list: TCR-induced genes\nContrast: WT+OVA vs. cKO+OVA', fgsea_Lck_lfc.res.WT.vs.KO.OVA, 9, 11)
plot(GSEA.WTKO_act_250_top)

GSEA.WTKO_act_250_top.blank <- GSEA.WTKO_act_250_top +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.WTKO_act_250_top.blank)
```

```{r}
tiff('Main_figures/Full/Lck_bulk_GSEA_WTKO_Act_Top_250_act.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_250_top)
dev.off()

tiff('Main_figures/Blank/Lck_bulk_GSEA_WTKO_Act_Top_250_act_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_250_top.blank)
dev.off()
```

GSEA WT/KO Act. contrast (NFAT-induced genes)

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list[['NFAT_up']],  lfc.res.WT.vs.KO.OVA)$data
GSEA.WTKO_act_NFAT_up <- plotGSEA(GSE_data,  'Gene list: NFAT-induced genes\nContrast: WT+OVA vs. cKO+OVA', fgsea_Lck_lfc.res.WT.vs.KO.OVA, 4, 11)
plot(GSEA.WTKO_act_NFAT_up)

GSEA.WTKO_act_NFAT_up.blank <- GSEA.WTKO_act_NFAT_up +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.WTKO_act_NFAT_up.blank)
```

```{r}
tiff('Main_figures/Full/Lck_bulk_GSEA_WTKO_Act_NFAT_up.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_NFAT_up)
dev.off()

tiff('Main_figures/Blank/Lck_bulk_GSEA_WTKO_Act_NFAT_up_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_NFAT_up.blank)
dev.off()
```

GSEA WT/KO Act. contrast (ERK-induced genes)

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list[['ERK_up']],  lfc.res.WT.vs.KO.OVA)$data
GSEA.WTKO_act_ERK_up <- plotGSEA(GSE_data,  'Gene list: ERK-induced genes\nContrast: WT+OVA vs. cKO+OVA', fgsea_Lck_lfc.res.WT.vs.KO.OVA, 2, 11)
plot(GSEA.WTKO_act_ERK_up)

GSEA.WTKO_act_ERK_up.blank <- GSEA.WTKO_act_ERK_up +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.WTKO_act_ERK_up.blank)
```

```{r}
tiff('Main_figures/Full/Lck_bulk_GSEA_WTKO_Act_ERK_up.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_ERK_up)
dev.off()

tiff('Main_figures/Blank/Lck_bulk_GSEA_WTKO_Act_ERK_up_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_ERK_up.blank)
dev.off()
```

GSEA WT/KO Act. contrast (P38-induced genes)

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list[['P38_up']], lfc.res.WT.vs.KO.OVA)$data
GSEA.WTKO_act_P38_up <- plotGSEA(GSE_data,  'Gene list: P38-induced genes\nContrast: WT+OVA vs. cKO+OVA', fgsea_Lck_lfc.res.WT.vs.KO.OVA, 7, 11)
plot(GSEA.WTKO_act_P38_up)

GSEA.WTKO_act_P38_up.blank <- GSEA.WTKO_act_P38_up +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.WTKO_act_P38_up.blank)
```

```{r}
tiff('Main_figures/Full/Lck_bulk_GSEA_WTKO_Act_P38_up.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_P38_up)
dev.off()

tiff('Main_figures/Blank/Lck_bulk_GSEA_WTKO_Act_P38_up_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_P38_up.blank)
dev.off()
```

GSEA WT/KO Act. contrast (NF-kB-induced genes)

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list[['NFKB_all']], lfc.res.WT.vs.KO.OVA)$data
GSEA.WTKO_act_NFKB_all <- plotGSEA(GSE_data, 'Gene list: NF-kB-induced genes\nContrast: WT+OVA vs. cKO+OVA', fgsea_Lck_lfc.res.WT.vs.KO.OVA, 5, 11)
plot(GSEA.WTKO_act_NFKB_all)

GSEA.WTKO_act_NFKB_all.blank <- GSEA.WTKO_act_NFKB_all +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.WTKO_act_NFKB_all.blank)
```

```{r}
tiff('Main_figures/Full/Lck_bulk_GSEA_WTKO_Act_NFKB_all.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_NFKB_all)
dev.off()

tiff('Main_figures/Blank/Lck_bulk_GSEA_WTKO_Act_NFKB_all_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_NFKB_all.blank)
dev.off()
```

## WT vs. KO - act - down

GSEA WT/KO Act. contrast (Activation-repressed genes)

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list[['top250act.down']],  lfc.res.WT.vs.KO.OVA)$data
GSEA.WTKO_act_250_top_inh <- plotGSEA(GSE_data, 'Gene list: TCR-repressed genes\nContrast: WT+OVA vs. cKO+OVA', fgsea_Lck_lfc.res.WT.vs.KO.OVA, 8, 11)
plot(GSEA.WTKO_act_250_top_inh)

GSEA.WTKO_act_250_top_inh.blank <- GSEA.WTKO_act_250_top_inh +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.WTKO_act_250_top_inh.blank)
```

```{r}
tiff('Supplemental_figures/Full/Lck_bulk_GSEA_WTKO_Act_Top_250_inh.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_250_top_inh)
dev.off()

tiff('Supplemental_figures/Blank/Lck_bulk_GSEA_WTKO_Act_Top_250_inh_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_250_top_inh.blank)
dev.off()
```

GSEA WT/KO Act. contrast (NFAT-repressed genes)

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list[['NFAT_down']], lfc.res.WT.vs.KO.OVA)$data
GSEA.WTKO_act_NFAT_down <- plotGSEA(GSE_data, 'Gene list: NFAT-repressed genes\nContrast: WT+OVA vs. cKO+OVA', fgsea_Lck_lfc.res.WT.vs.KO.OVA, 3, 11)
plot(GSEA.WTKO_act_NFAT_down)

GSEA.WTKO_act_NFAT_down.blank <- GSEA.WTKO_act_NFAT_down +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.WTKO_act_NFAT_down.blank)
```

```{r}
tiff('Supplemental_figures/Full/Lck_bulk_GSEA_WTKO_Act_NFAT_down.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_NFAT_down)
dev.off()

tiff('Supplemental_figures/Blank/Lck_bulk_GSEA_WTKO_Act_NFAT_down_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_NFAT_down.blank)
dev.off()
```

GSEA WT/KO Act. contrast (ERK-repressed genes)

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list[['ERK_down']], lfc.res.WT.vs.KO.OVA)$data
GSEA.WTKO_act_ERK_down <- plotGSEA(GSE_data, 'Gene list: ERK-repressed genes\nContrast: WT+OVA vs. cKO+OVA', fgsea_Lck_lfc.res.WT.vs.KO.OVA, 1, 11)
plot(GSEA.WTKO_act_ERK_down)

GSEA.WTKO_act_ERK_down.blank <- GSEA.WTKO_act_ERK_down +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.WTKO_act_ERK_down.blank)
```

```{r}
tiff('Supplemental_figures/Full/Lck_bulk_GSEA_WTKO_Act_ERK_down.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_ERK_down)
dev.off()

tiff('Supplemental_figures/Blank/Lck_bulk_GSEA_WTKO_Act_ERK_down_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_ERK_down.blank)
dev.off()
```

GSEA WT/KO Act. contrast (P38-repressed genes)

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list[['P38_down']], lfc.res.WT.vs.KO.OVA)$data
GSEA.WTKO_act_P38_down <- plotGSEA(GSE_data, 'Gene list: P38-repressed genes\nContrast: WT+OVA vs. cKO+OVA', fgsea_Lck_lfc.res.WT.vs.KO.OVA, 6, 11)
plot(GSEA.WTKO_act_P38_down)

GSEA.WTKO_act_P38_down.blank <- GSEA.WTKO_act_P38_down +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.WTKO_act_P38_down.blank)
```

```{r}
tiff('Supplemental_figures/Full/Lck_bulk_GSEA_WTKO_Act_P38_down.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_P38_down)
dev.off()

tiff('Supplemental_figures/Blank/Lck_bulk_GSEA_WTKO_Act_P38_down_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_P38_down.blank)
dev.off()
```

## WT inhibitors comparisons.

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list[['NFAT_up']],  lfc.res.WT.vs.WT.NFAT)$data
GSEA.WT_act_vs_act_Cycl_NFAT_up <- plotGSEA(GSE_data, 'Gene list: NFAT-induced genes\nContrast: WT+OVA vs. WT+OVA+NFATi', fgsea_Lck_lfc.res.WT.vs.WT.NFAT, 4, 11)
plot(GSEA.WT_act_vs_act_Cycl_NFAT_up)

GSEA.WT_act_vs_act_Cycl_NFAT_up.blank <- GSEA.WT_act_vs_act_Cycl_NFAT_up +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.WT_act_vs_act_Cycl_NFAT_up.blank)
```

```{r}
tiff('Supplemental_figures/Full/Lck_bulk_GSEA_WT_Act_vs_act_NFAT_Cycl_up.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WT_act_vs_act_Cycl_NFAT_up)
dev.off()

tiff('Supplemental_figures/Blank/Lck_bulk_GSEA_WT_Act_vs_act_NFAT_Cycl_up_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WT_act_vs_act_Cycl_NFAT_up.blank)
dev.off()
```

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list[['ERK_up']],  lfc.res.WT.vs.WT.ERK)$data
GSEA.WT_act_vs_act_ERK_ERK_up <- plotGSEA(GSE_data, 'Gene list: ERK-induced genes\nContrast: WT+OVA vs. WT+OVA+ERKi', fgsea_Lck_lfc.res.WT.vs.WT.ERK, 2, 11)
plot(GSEA.WT_act_vs_act_ERK_ERK_up)

GSEA.WT_act_vs_act_ERK_ERK_up.blank <- GSEA.WT_act_vs_act_ERK_ERK_up +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.WT_act_vs_act_ERK_ERK_up.blank)
```

```{r}
tiff('Supplemental_figures/Full/Lck_bulk_GSEA_WT_Act_vs_Act_ERK_ERK_up.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WT_act_vs_act_ERK_ERK_up)
dev.off()

tiff('Supplemental_figures/Blank/Lck_bulk_GSEA_WT_Act_vs_Act_ERK_ERK_up_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WT_act_vs_act_ERK_ERK_up.blank)
dev.off()
```

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list[['P38_up']],  lfc.res.WT.vs.WT.IP38)$data
GSEA.WT_act_vs_act_Ip38_P38_up <- plotGSEA(GSE_data, 'Gene list: P38-induced genes\nContrast: WT+OVA vs. WT+OVA+P38i', fgsea_Lck_lfc.res.WT.vs.WT.IP38, 7, 11)
plot(GSEA.WT_act_vs_act_Ip38_P38_up)

GSEA.WT_act_vs_act_Ip38_P38_up.blank <- GSEA.WT_act_vs_act_Ip38_P38_up +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.WT_act_vs_act_Ip38_P38_up.blank)
```

```{r}
tiff('Supplemental_figures/Full/Lck_bulk_GSEA_WT_Act_vs_Act_Ip38_P38_up.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WT_act_vs_act_Ip38_P38_up)
dev.off()

tiff('Supplemental_figures/Blank/Lck_bulk_GSEA_WT_Act_vs_Act_Ip38_P38_up_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WT_act_vs_act_Ip38_P38_up.blank)
dev.off()
```

## KO inhibitors comparisons.

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list[['NFAT_up']],  lfc.res.KO.vs.KO.NFAT)$data
GSEA.KO_act_vs_act_Cycl_NFAT_up <- plotGSEA(GSE_data, 'Gene list: NFAT-induced genes\nContrast: cKO+OVA vs. cKO+OVA+NFATi', fgsea_Lck_lfc.res.KO.vs.KO.NFAT, 4, 11)
plot(GSEA.KO_act_vs_act_Cycl_NFAT_up)

GSEA.KO_act_vs_act_Cycl_NFAT_up.blank <- GSEA.KO_act_vs_act_Cycl_NFAT_up +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.KO_act_vs_act_Cycl_NFAT_up.blank)
```

```{r}
tiff('Supplemental_figures/Full/Lck_bulk_GSEA_KO_Act_vs_act_NFAT_Cycl_up.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.KO_act_vs_act_Cycl_NFAT_up)
dev.off()

tiff('Supplemental_figures/Blank/Lck_bulk_GSEA_KO_Act_vs_act_NFAT_Cycl_up_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.KO_act_vs_act_Cycl_NFAT_up.blank)
dev.off()
```

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list[['ERK_up']],  lfc.res.KO.vs.KO.ERK)$data
GSEA.KO_act_vs_act_ERK_ERK_up <- plotGSEA(GSE_data, 'Gene list: ERK-induced genes\nContrast: cKO+OVA vs. cKO+OVA+ERKi', fgsea_Lck_lfc.res.KO.vs.KO.ERK, 2, 11)
plot(GSEA.KO_act_vs_act_ERK_ERK_up)

GSEA.KO_act_vs_act_ERK_ERK_up.blank <- GSEA.KO_act_vs_act_ERK_ERK_up +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.KO_act_vs_act_ERK_ERK_up.blank)
```

```{r}
tiff('Supplemental_figures/Full/Lck_bulk_GSEA_KO_Act_vs_Act_ERK_ERK_up.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.KO_act_vs_act_ERK_ERK_up)
dev.off()

tiff('Supplemental_figures/Blank/Lck_bulk_GSEA_KO_Act_vs_Act_ERK_ERK_up_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.KO_act_vs_act_ERK_ERK_up.blank)
dev.off()
```

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list[['P38_up']],  lfc.res.KO.vs.KO.IP38)$data
GSEA.KO_act_vs_act_Ip38_P38_up <- plotGSEA(GSE_data, 'Gene list: P38-induced genes\nContrast: cKO+OVA vs. cKO+OVA+P38i', fgsea_Lck_lfc.res.KO.vs.KO.IP38, 7, 11)
plot(GSEA.KO_act_vs_act_Ip38_P38_up)

GSEA.KO_act_vs_act_Ip38_P38_up.blank <- GSEA.KO_act_vs_act_Ip38_P38_up +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.KO_act_vs_act_Ip38_P38_up.blank)
```

```{r}
tiff('Supplemental_figures/Full/Lck_bulk_GSEA_KO_Act_vs_Act_Ip38_P38_up.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.KO_act_vs_act_Ip38_P38_up)
dev.off()

tiff('Supplemental_figures/Blank/Lck_bulk_GSEA_KO_Act_vs_Act_Ip38_P38_up_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.KO_act_vs_act_Ip38_P38_up.blank)
dev.off()
```

## WT/KO inhibitor comparisons

GSEA WT/KO Cyclosp. A contrast (Activation-induced genes)

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list[['top250act.up']],  lfc.res.WT.NFAT.vs.KO.NFAT)$data
GSEA.WTKO_act_cyclosp_250_top <- plotGSEA(GSE_data, 'Gene list: TCR-induced genes\nContrast: WT+OVA+NFATi vs. cKO+OVA+NFATi', fgsea_Lck_lfc.res.WT.NFAT.vs.KO.NFAT, 9, 11)
plot(GSEA.WTKO_act_cyclosp_250_top)

GSEA.WTKO_act_cyclosp_250_top.blank <- GSEA.WTKO_act_cyclosp_250_top  +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.WTKO_act_cyclosp_250_top.blank)
```

```{r}
tiff('Main_figures/Full/Lck_bulk_GSEA_WTKO_Act_Cyclosp_A_Top_250_act.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_cyclosp_250_top)
dev.off()

tiff('Main_figures/Blank/Lck_bulk_GSEA_WTKO_Act_Cyclosp_A_Top_250_act_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_cyclosp_250_top.blank)
dev.off()
```

GSEA WT/KO Act. + ERK contrast (Activation-induced genes)

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list[['top250act.up']],  lfc.res.WT.ERK.vs.KO.ERK)$data
GSEA.WTKO_act_ERK_250_top <- plotGSEA(GSE_data, 'Gene list: TCR-induced genes\nContrast: WT+OVA+ERKi vs. cKO+OVA+ERKi', fgsea_Lck_lfc.res.WT.ERK.vs.KO.ERK, 9, 11)
plot(GSEA.WTKO_act_ERK_250_top)

GSEA.WTKO_act_ERK_250_top.blank <- GSEA.WTKO_act_ERK_250_top  +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.WTKO_act_ERK_250_top.blank)
```

```{r}
tiff('Main_figures/Full/Lck_bulk_GSEA_WTKO_Act_ERK_A_Top_250_act.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_ERK_250_top)
dev.off()

tiff('Main_figures/Blank/Lck_bulk_GSEA_WTKO_Act_ERK_A_Top_250_act_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_ERK_250_top.blank)
dev.off()
```

GSEA WT/KO Act. + P38 contrast (Activation-induced genes)

```{r, fig.width = 9, fig.show= 'hold', fig.align = 'center'}
GSE_data <- plotEnrichment(pathways_list[['top250act.up']],  lfc.res.WT.IP38.vs.KO.IP38)$data
GSEA.WTKO_act_P38_250_top <- plotGSEA(GSE_data, 'Gene list: TCR-induced genes\nContrast: WT+OVA+P38i vs. cKO+OVA+P38i', fgsea_Lck_lfc.res.WT.IP38.vs.KO.IP38, 9, 11)
plot(GSEA.WTKO_act_P38_250_top)

GSEA.WTKO_act_P38_250_top.blank <- GSEA.WTKO_act_P38_250_top  +
  theme(text = element_text(color = "white"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  strip.text = element_text(color = "white"),
  axis.text.x = element_text(color = "white"),
  axis.text.y = element_text(color = "white")
)

print(GSEA.WTKO_act_P38_250_top.blank)
```

```{r}
tiff('Main_figures/Full/Lck_bulk_GSEA_WTKO_Act_P38_A_Top_250_act.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_P38_250_top)
dev.off()

tiff('Main_figures/Blank/Lck_bulk_GSEA_WTKO_Act_P38_A_Top_250_act_blank.tiff', width = 1700, height = 1000, res = 300)
print(GSEA.WTKO_act_P38_250_top.blank)
dev.off()
```

# Heatmap of L2FC for each sample

We calculate L2FC for each meaningful contrast between WT and KO samples for each sample separately. First, show colData of DDS object.

```{r}
print(dds_bulk_Lck.data@colData %>% as.data.frame)
```

Now compute fold changes for each contrast and sample. This will be computed from normalized data.

```{r}
l2fc.WT.vs.KO.sample.2 <- log2((Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample6Bulk1"] + 1)/(Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample11Bulk1"] + 1))
l2fc.WT.vs.KO.sample.3 <- log2((Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample1Bulk2"] + 1)/(Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample6Bulk2"] + 1))

l2fc.WT.vs.KO.OVA.sample.2 <- log2((Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample7Bulk1"] + 1)/(Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample12Bulk1"] + 1))
l2fc.WT.vs.KO.OVA.sample.3 <- log2((Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample2Bulk2"] + 1)/(Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample7Bulk2"] + 1))

l2fc.WT.vs.KO.p38i.sample.2 <- log2((Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample8Bulk1"] + 1)/(Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample13Bulk1"] + 1))
l2fc.WT.vs.KO.p38i.sample.3 <- log2((Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample3Bulk2"] + 1)/(Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample8Bulk2"] + 1))

l2fc.WT.vs.KO.ERKi.sample.2 <- log2((Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample9Bulk1"] + 1)/(Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample14Bulk1"] + 1))
l2fc.WT.vs.KO.ERKi.sample.3 <- log2((Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample4Bulk2"] + 1)/(Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample9Bulk2"] + 1))

l2fc.WT.vs.KO.NFATi.sample.2 <- log2((Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample10Bulk1"] + 1)/(Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample15Bulk1"] + 1))
l2fc.WT.vs.KO.NFATi.sample.3 <- log2((Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample5Bulk2"] + 1)/(Lck_normalized[,dds_bulk_Lck.data$sample.name == "Sample10Bulk2"] + 1))
```

Build matrix from the data and name it.

```{r}
lf2c.matrix.sample <- cbind(l2fc.WT.vs.KO.sample.2, l2fc.WT.vs.KO.sample.3,
                            l2fc.WT.vs.KO.OVA.sample.2, l2fc.WT.vs.KO.OVA.sample.3,
                            l2fc.WT.vs.KO.p38i.sample.2, l2fc.WT.vs.KO.p38i.sample.3,
                            l2fc.WT.vs.KO.ERKi.sample.2, l2fc.WT.vs.KO.ERKi.sample.3,
                            l2fc.WT.vs.KO.NFATi.sample.2, l2fc.WT.vs.KO.NFATi.sample.3) %>% as.matrix

rownames(lf2c.matrix.sample) <- names(l2fc.WT.vs.KO.sample.2)
colnames(lf2c.matrix.sample) <- c('L2FC-WT/KO-NA.bulk1', 'L2FC-WT/KO-NA.bulk2',
                                  'L2FC-WT/KO-OVA.bulk1', 'L2FC-WT/KO-OVA.bulk2',
                                  'L2FC-WT/KO-p38i.bulk1', 'L2FC-WT/KO-p38i.bulk2',
                                  'L2FC-WT/KO-ERKi.bulk1', 'L2FC-WT/KO-ERKi.bulk2',
                                  'L2FC-WT/KO-NFATi.bulk1', 'L2FC-WT/KO-NFATi.bulk2')
```

Make some validations.

```{r}
test_bulk1 <- log2((Lck_normalized[,1:5] + 1)/(Lck_normalized[,6:10] + 1))
test_bulk2 <- log2((Lck_normalized[,11:15] + 1)/(Lck_normalized[,16:20] + 1))

print(test_bulk1 %>% head)
print(test_bulk2 %>% head)

do.call(
  rbind,
  lapply(gene_list_files, function(f) calc_gene_score(f, test_bulk1, do.scaling = FALSE))
)


do.call(
  rbind,
  lapply(gene_list_files, function(f) calc_gene_score(f, test_bulk2, do.scaling = FALSE))
)

```

Results are identical with above matrix, so that's nice.

Paths of files was defined before using parameter 'folder' and 'gene_list_files'. We use scoring function without scaling to obtain L2FC.

```{r}
# calculate scores for all gene lists
L2FCScoreTable <- do.call(
  rbind,
  lapply(gene_list_files, function(f) calc_gene_score(f, lf2c.matrix.sample, do.scaling = FALSE))
)

rownames(L2FCScoreTable) <- tools::file_path_sans_ext(basename(gene_list_files))
```

Save the score table.

```{r}
write.csv(L2FCScoreTable, file.path(folder, "L2FC_ScoreTable.csv"), quote = FALSE)
```

Plot heatmaps.

```{r}
#HEAT MAP WITH ALL 10 SAMPLES
pheatmap(
  as.matrix(L2FCScoreTable),
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  color = colorRampPalette(c("blue", "white", "red"))(100)
)

#HEAT MAP WITH 8 SAMPLES WHICH WERE OVA-ACTIVATED
pheatmap(
  as.matrix(L2FCScoreTable[, -c(1:2)]),
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  color = colorRampPalette(c("blue", "white", "red"))(100)
)

#WITHOUT inhibitors
pheatmap(
  as.matrix(L2FCScoreTable[, c(1:4)]),
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  color = colorRampPalette(c("blue", "white", "red"))(100)
)

#ONLY SAMPLES WITHOUT Inhibitors - COLOR SCALING
mat <- as.matrix(L2FCScoreTable[, c(1:4)])

my_breaks <- c(
  seq(-2, 0, length.out = 20),
  seq(0.001, 1.5, length.out = 80)
)

#HEAT MAP WITH 8 SAMPLES WHICH WERE OVA-ACTIVATED
pheatmap(
  as.matrix(L2FCScoreTable[, -c(1:4)]),
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  color = colorRampPalette(c("blue", "white", "red"))(100),
  breaks = my_breaks
)

pheatmap(
  mat,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  color = colorRampPalette(c("blue", "white", "red"))(100),
  breaks = my_breaks
)
```

```{r}
lim <- max(abs(range(L2FCScoreTable[, c(1:4)], na.rm = TRUE)))

cols <- colorRampPalette(c("blue", "white", "red"))(101)
breaks <- seq(-lim, lim, length.out = 102)

pheatmap(L2FCScoreTable[, c(1:4)],
         cluster_rows = FALSE,
         cluster_cols = FALSE,         
         color = cols,
         breaks = breaks)

#lim <- max(abs(range(L2FCScoreTable[, -c(1:2)], na.rm = TRUE)))

#cols <- colorRampPalette(c("blue", "white", "red"))(101)
#breaks <- seq(-lim, lim, length.out = 102)

pheatmap(L2FCScoreTable[, -c(1:2)],
         cluster_rows = FALSE,
         cluster_cols = FALSE,         
         color = cols,
         breaks = breaks)
```

```{r}
lim <- max(abs(range(L2FCScoreTable[c(1,4,9), -c(1:2)], na.rm = TRUE)))

cols <- colorRampPalette(c("blue", "white", "red"))(101)
breaks <- seq(-lim, lim, length.out = 102)

pheatmap(L2FCScoreTable[c(1,4,9), -c(1:2)],
         cluster_rows = FALSE,
         cluster_cols = FALSE,         
         color = cols,
         breaks = breaks)
```

```{r}
plot.mat.act <- cbind(L2FCScoreTable[c(1), -c(1:2,4,6,8,10)], L2FCScoreTable[c(1), -c(1:2,3,5,7,9)]) 
rownames(plot.mat.act) <- c('L2FC-WT/KO-OVA', 'L2FC-WT/KO-p38i', 'L2FC-WT/KO-ERKi', 'L2FC-WT/KO-NFATi')
colnames(plot.mat.act) <- c('Replicate1', 'Replicate2')

lim <- max(abs(range(plot.mat.act, na.rm = TRUE)))

cols <- colorRampPalette(c("blue", "white", "red"))(101)
breaks <- seq(-lim, lim, length.out = 102)

pheatmap(plot.mat.act,
         cluster_rows = FALSE,
         cluster_cols = FALSE,         
         color = cols,
         breaks = breaks)
```
