---
title: "Experiment LCK"
author: "Juraj Michalik"
date: "15/7/2025"
output: rmdformats::material
---

This script is used to import supplementary assays to LCK data Seurat object after second processing step.

# Initialization 

Load all necessary packages here.

```{r libs, include=FALSE}
knitr::opts_chunk$set(include = TRUE, warning = FALSE, message = FALSE, error = TRUE, cache = TRUE)
library(Seurat)
library(future)
library(ggplot2)
library(dplyr)
library(tidyr)
library(cowplot)
library(gridExtra)
library(dbscan)
library(scatterpie)
library(reshape2)
library(ggrepel)
library(scds)
library(SingleCellExperiment)
library(homologene)
library(gprofiler2)
library(STACAS)
library(SingleR)
library(ComplexHeatmap)

source('VDJ_utilities.R')
source('VDJ_TRUST4_import.R')

# negation of %in% 
`%nin%` = Negate(`%in%`)
```

Get cell cycle genes.

```{r}
mm.s.genes <- scGate::genes.blacklist.default$Mm$cellCycle.G1S
mm.g2m.genes <- scGate::genes.blacklist.default$Mm$cellCycle.G2M
```

# Analyses of Lck vs. Lck WT correlation

Load processed data set.

```{r}
exp43_Lck_filt <- readRDS('Datasets/exp43_Lck_processed_new_with_vdj.rds')
```

```{r}
raw_exp43_Lck <- GetAssayData(exp43_Lck_filt, assay = 'RNA', slot = 'counts')
raw_exp43_LckWT <- GetAssayData(exp43_Lck_filt, assay = 'WTKO', slot = 'counts')

Lck.df <- data.frame(Lck = raw_exp43_Lck['Lck',], LckWT = raw_exp43_LckWT['Lck',], Condition =  exp43_Lck_filt$Condition) 
```

Plots.

```{r}
Lck.df.WT <- Lck.df[Lck.df$Condition == 'WT',]
Lck.df.KO <- Lck.df[Lck.df$Condition == 'cLCK KO',]

ggplot(Lck.df.WT, aes(x = Lck, y = LckWT)) + geom_point() + geom_density_2d_filled(bins = 6, alpha = 0.5) + ggtitle('Lck vs. LckWT, WT mice') + theme_classic() + xlim(0, 32) + ylim(0, 32)
ggplot(Lck.df.KO, aes(x = Lck, y = LckWT)) + geom_point() + geom_density_2d_filled(bins = 6, alpha = 0.5) + ggtitle('Lck vs. LckWT, cLCK KO mice') + theme_classic() + xlim(0, 32) + ylim(0, 32)

Lck.df.WT.counted <- Lck.df.WT %>% dplyr::count(Lck, LckWT)
Lck.df.KO.counted <- Lck.df.KO %>% dplyr::count(Lck, LckWT)

ggplot(Lck.df.WT.counted, aes(x = Lck, y = LckWT, color = n)) + geom_point() + ggtitle('Lck vs. LckWT, WT mice') + theme_classic() + xlim(0, 32) + ylim(0, 32) + scale_color_gradient2(low = 'blue', mid = 'gray85', high = 'firebrick3')
ggplot(Lck.df.KO.counted, aes(x = Lck, y = LckWT, color = n)) + geom_point() + ggtitle('Lck vs. LckWT, cLCK KO mice') + theme_classic() + xlim(0, 32) + ylim(0, 32) + scale_color_gradient2(low = 'blue', mid = 'gray85', high = 'firebrick3')
```


Make more advanced table.

```{r}
raw_exp43_Lck <- GetAssayData(exp43_Lck_filt, assay = 'RNA', slot = 'counts')
raw_exp43_LckWT <- GetAssayData(exp43_Lck_filt, assay = 'WTKO', slot = 'counts')
Lck.df.meta <- data.frame(Lck = raw_exp43_Lck['Lck',], LckWT = raw_exp43_LckWT['Lck',])
Lck.df.meta <- cbind(Lck.df.meta, exp43_Lck_filt@meta.data) 

Lck.df.meta.filt <- Lck.df.meta %>% dplyr::select(Lck, LckWT, Mouse_ID) %>%
  dplyr::filter(LckWT == 0) 

mouse_info <- Lck.df.meta %>%
  distinct(Mouse_ID, Condition)

Lck.df.LckWT.null <- Lck.df.meta.filt %>% group_by(Lck, Mouse_ID) %>%
  dplyr::summarise(count = n()) %>% ungroup() %>% 
  tidyr::complete(Lck, Mouse_ID, fill = list(count = 0)) %>%
  dplyr::left_join(mouse_info, by = "Mouse_ID") %>% 
  dplyr::group_by(Lck) %>%
  dplyr::arrange(Condition, Mouse_ID, .by_group = T)

write.csv(Lck.df.LckWT.null, 'Lck_null_mouse_counts.csv')
```

Import Cd45 data.

```{r}
Cd45_assay <- Read10X_h5('Lck_Cd45_genotyping/raw_feature_bc_matrix.h5')
Cd45_assay <- Cd45_assay$`Gene Expression`[c('Cd45.1', 'Cd45.2'),]
```

```{r}
Cd45_assay <- Cd45_assay[,colnames(Cd45_assay) %in% colnames(exp43_Lck_filt)]
```

```{r}
exp43_Lck_filt[['CD45']] <- CreateAssayObject(Cd45_assay)
```

```{r}
FeaturePlot(exp43_Lck_filt, c('Cd45.1', 'Cd45.2'))
```

```{r}
saveRDS(exp43_Lck_filt, 'Datasets/exp43_Lck_processed_new_with_vdj_and_cd45.rds')
```

# Heatmaps by entites

We make heatmaps by entites (mice split by whether they have Lck aligned to them). First, we determine whether cells are Lck+ or Lck- based on WTKO assay.

```{r}
LckWT.counts <- GetAssayData(exp43_Lck_filt, assay = 'WTKO', slot = 'counts')
exp43_Lck_filt$Lck.exons <- ifelse(LckWT.counts['Lck',] > 0, 'Lck+', 'Lck-')
exp43_Lck_filt$Mouse_ID.Lck.exons <- paste(exp43_Lck_filt$Mouse_ID, exp43_Lck_filt$Lck.exons, sep = '.')
```

Now make pseudobulk based on these data.

```{r}
# Make unique identifier 

# Make meta-data
exp43.Lck.bulk.meta <- exp43_Lck_filt@meta.data %>% dplyr::select(Mouse_ID.Lck.exons, Mouse_ID, Lck.exons, Condition) %>% 
  .[!duplicated(.),] %>% dplyr::arrange(Condition, Lck.exons, Mouse_ID)
rownames(exp43.Lck.bulk.meta) <- exp43.Lck.bulk.meta$Mouse_ID.Lck.exons
exp43.Lck.bulk.meta <- exp43.Lck.bulk.meta %>% dplyr::select(-Mouse_ID.Lck.exons)

# Get data and arrange them by cell type
exp43.Lck.counts <- AggregateExpression(exp43_Lck_filt, group.by = 'Mouse_ID.Lck.exons', slot = 'counts', assays = 'RNA')
colnames(exp43.Lck.counts$RNA) <- gsub('g', '', colnames(exp43.Lck.counts$RNA))
exp43.Lck.counts$RNA <- exp43.Lck.counts$RNA[,match(rownames(exp43.Lck.bulk.meta), colnames(exp43.Lck.counts$RNA))]
exp43.Lck.pbulk <- CreateSeuratObject(exp43.Lck.counts, project = 'E43 Lck - entities')
exp43.Lck.pbulk <- NormalizeData(exp43.Lck.pbulk, scale.factor = 10^6)

# Add Meta data
exp43.Lck.pbulk <- AddMetaData(exp43.Lck.pbulk, exp43.Lck.bulk.meta)

# Scale and calculate PCA for later
exp43.Lck.pbulk <- ScaleData(exp43.Lck.pbulk, features = rownames(exp43.Lck.pbulk))
exp43.Lck.pbulk <- RunPCA(exp43.Lck.pbulk, npcs = 3, features = rownames(exp43.Lck.pbulk))
```

Perform DEA on cLck KO vs WT.

```{r}
Idents(exp43_Lck_filt) <- 'Condition'
exp43_Lck_filt.DEA <- FindMarkers(exp43_Lck_filt, assay = 'RNA', ident.1 = 'cLCK KO', ident.2 = 'WT')
Idents(exp43_Lck_filt) <- 'seurat_clusters'

write.csv(exp43_Lck_filt.DEA %>% dplyr::filter(p_val_adj < 0.05 & abs(avg_log2FC) > 0.25), 'DEA/exp43_Lck_filt_DEA_cLCKO_WT.csv')
```

Show DEA.

```{r}
print(head(exp43_Lck_filt.DEA %>% dplyr::filter(abs(avg_log2FC) > 0.25), n = 100))
```

Plot heatmap of top 100 genes.

```{r, fig.width=7, fig.height=10}
# Make blank heatmap
colfunc <- colorRampPalette(c('blue', "gray90", 'firebrick2'))

# Assemble matrix
assay_data <- GetAssayData(exp43.Lck.pbulk)

exp43.Lck.bulk.meta <- exp43.Lck.bulk.meta[c('Condition', 'Lck.exons')]
colors.condition <- c('steelblue', 'firebrick1')
names(colors.condition) <- c('WT', 'cLCK KO')
colors.exons <- c('dodgerblue', 'firebrick')
names(colors.exons) <- c('Lck+', 'Lck-')

ann_colors <- list(colors.condition, colors.exons)
names(ann_colors) <- c('Condition', 'Lck.exons')
col_annot <- HeatmapAnnotation(df = exp43.Lck.bulk.meta, col = ann_colors,
                               annotation_legend_param = list(
                                 title_gp = gpar(col = 'black', fontface = 'bold'),
                                 labels_gp = gpar(col = 'black')
                                ))

Lck.Top50.hmap <- ComplexHeatmap::Heatmap(assay_data[rownames(exp43_Lck_filt.DEA)[1:50],] %>% as.matrix %>% t %>% scale %>% t,
                                           name = 'Z-score', color=colfunc(64), cluster_columns = F, border = NA,
                                           top_annotation = col_annot,
                                           row_title_rot = 0,
                                           column_title = "Heatmap of selected genes from\nLck Murine experiment, by entities",
                                           column_title_gp = gpar(fontface = "bold", col = 'white'),
                                           row_names_gp = gpar(col = "black"),
                                           row_title_gp = gpar(col = "black"),
                                           column_names_gp = gpar(col = "black"),
                                           heatmap_legend_param = list(
                                             title_gp = gpar(col = 'black', fontface = 'bold'),
                                             labels_gp = gpar(col = 'black')
                                           ))

print(Lck.Top50.hmap)
```

```{r, fig.width=5, fig.height=5}
# Make blank heatmap
colfunc <- colorRampPalette(c('blue', "gray90", 'firebrick2'))

# Assemble matrix
assay_data <- GetAssayData(exp43.Lck.pbulk)

exp43.Lck.bulk.meta <- exp43.Lck.bulk.meta[c('Condition', 'Lck.exons')]
colors.condition <- c('steelblue', 'firebrick1')
names(colors.condition) <- c('WT', 'cLCK KO')
colors.exons <- c('dodgerblue', 'firebrick')
names(colors.exons) <- c('Lck+', 'Lck-')

ann_colors <- list(colors.condition, colors.exons)
names(ann_colors) <- c('Condition', 'Lck.exons')
col_annot <- HeatmapAnnotation(df = exp43.Lck.bulk.meta, col = ann_colors,
                               annotation_legend_param = list(
                                 title_gp = gpar(col = 'black', fontface = 'bold'),
                                 labels_gp = gpar(col = 'black')
                                ))

Lck.eff.hmap <- ComplexHeatmap::Heatmap(assay_data[c('Lck', 'Gzma', 'Gzmb', 'Klrg1', 'Klrk1', 'Ccl5', 'Trgc4', 'Trdc', 'Klrd1', 'Klre1', 'Klrc1', 'Id2', 'Isg15', 'Gzmm', 'Lrrk1', 'Serpinb6a', 'Cxcr6'),] %>% as.matrix %>% t %>% scale %>% t,
                                        name = 'Z-score', color=colfunc(64), cluster_columns = F, border = NA,
                                        top_annotation = col_annot,
                                        row_title_rot = 0,
                                        column_title = "Heatmap of selected genes from\nLck Murine experiment, by entities",
                                        column_title_gp = gpar(fontface = "bold", col = 'white'),
                                        row_names_gp = gpar(col = "black"),
                                        row_title_gp = gpar(col = "black"),
                                        column_names_gp = gpar(col = "black"),
                                        heatmap_legend_param = list(
                                          title_gp = gpar(col = 'black', fontface = 'bold'),
                                          labels_gp = gpar(col = 'black')
                                          ))

print(Lck.eff.hmap)
```

Filter out strange cells:
- Those that have cdr3a that is different from CAASDNYQLIW or those with cdr3b different CASSRANYEQYF, or any second cdr3a chain
- Those that have CD45.1
- Those with 0 Lck raw counts

First mark cells by whit OT-I they possess

```{r}
exp43_Lck_finalized <- exp43_Lck_filt 
exp43_Lck_finalized$OT.I <- 'OT-I not disproved'
exp43_Lck_finalized$OT.I[!(is.na(exp43_Lck_finalized$cdr3_B) | (adist(exp43_Lck_finalized$cdr3_B, 'CASSRANYEQYF') < 2))] <- 'Non OT-I'
exp43_Lck_finalized$OT.I[!(is.na(exp43_Lck_finalized$cdr3_A1) | (adist(exp43_Lck_finalized$cdr3_A1, 'CAASDNYQLIW') < 2))] <- 'Marked'
exp43_Lck_finalized$OT.I[exp43_Lck_finalized$OT.I == 'Marked' & !is.na(exp43_Lck_finalized$cdr3_A2) & adist(exp43_Lck_finalized$cdr3_A2, 'CAASDNYQLIW') < 2] <- 'OT-I not disproved'
exp43_Lck_finalized$OT.I[exp43_Lck_finalized$OT.I == 'Marked'] <- 'Non OT-I'
```

Mark cells by Cd45 expression and by Lck expression.

```{r}
Cd45.counts <- GetAssayData(exp43_Lck_finalized, assay = 'CD45', slot = 'counts')['Cd45.1',]
gene.counts <- GetAssayData(exp43_Lck_finalized, assay = 'RNA', slot = 'counts')['Lck',]

exp43_Lck_finalized$Cd45.count <- Cd45.counts
exp43_Lck_finalized$Lck.full.count <- gene.counts 
```

Filtering.

```{r}
exp43_Lck_finalized <- subset(exp43_Lck_finalized, OT.I == 'OT-I not disproved' & Cd45.count == 0 & Lck.full.count > 0)
```

```{r}
saveRDS(exp43_Lck_finalized, 'Datasets/exp43_Lck_processed_filt_Lck_OTI_Ly51.rds')
```

